---
title: "Cascade of Care CUD"
author: "Luis Segura"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
    highlight: kate
    self-contained: true
    df_print: paged
---

```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  fig.align = "center",
  fig.width = 7,
  fig.height = 8,
  out.width = "90%",
  cache = TRUE
)

### load packages
pckgs <- c("tidyverse", "here", "survey", "srvyr")

for (package in pckgs) {
  library(package, character.only = T)
}

```


```{r, echo = F}
## Loading dataset
load(here("data/", "nsduh_4analysis.Rdata"))

```

First, create the survey design object. I am also recoding some variables. Note that I am combining four binary variables to create the variable for healthcare utilization `hlth_util`. These variables are:

- `NMERTMT2` asks: During the past 12 months, that is since DATEFILL, how many different times have you been treated in an emergency room for any reason?

- `INHOSPYR` asks: During the past 12 months, have you stayed overnight or longer as an inpatient in a hospital? This is a binary no/yes variable that seems to measure hospitalization.

- `NMVSOPT2` asks: During the past 12 months, how many times have you visited a doctor, nurse, physician assistant or nurse practitioner about your own health at a doctor’s office, a clinic, or some other place?

- `TELEAPTYR2` asks: During the past 12 months, have you talked to a doctor, physician assistant, or nurse practitioner about your own health over the phone, by email, or through video calling instead of going to an in-person appointment? 

We also created a combined binary variable for CUD outpatient treatment `cud_trtment` from:

- `SUTOUTMRJPY` asks: RCVD SUB USE TRMT AS AN OUTPATIENT FOR MARIJUANA USE - PY. This is a binary variable yes/no

- `SUTOUTUNSPPY` asks: RCVD SUB USE TRMT AS AN OUTPATIENT - SUB UNSPECIFIED - PY

```{r, echo = F}
## creating survey design object and some recoding
nsduh_design <- nsduh_data |>
  mutate(cud = factor(irpyud5mrj, 
                              labels = c("No CUD", "Yes CUD")), 
         nmertmt2 = factor(ifelse(nmertmt2 > 900, NA, 
                           ifelse(nmertmt2 > 0 & nmertmt2 < 33, 1, nmertmt2)), 
                           labels = c("No ED visit", "ED visit")),
         inhospyr = factor(ifelse(inhospyr > 80, NA, 
                           ifelse(inhospyr == 2 , 0, inhospyr)), 
                           labels = c("No hospitalization", "Yes hospitalization")),
         nmvsopt2 = factor(ifelse(nmvsopt2 > 900, NA, 
                                  ifelse(nmvsopt2 > 0 & nmvsopt2 < 20 , 1, nmvsopt2)), 
                           labels = c("No outpatient visit", "Yes outpatient visit")),
         teleaptyr2 = factor(teleaptyr2, 
                             labels = c("No telemedicine", "Yes telemedicine")),
         hlth_util = factor(ifelse(nmertmt2 == "ED visit" | inhospyr == "Yes hospitalization" | 
                                     nmvsopt2 == "Yes outpatient visit" | teleaptyr2 == "Yes telemedicine", 1, 0), 
                            labels = c("No", "Yes Healthcare Utilization")), 
         hpusedrg = factor(ifelse(hpusedrg > 80, NA, 
                                  ifelse(hpusedrg == 2, 0, hpusedrg)), 
                           labels = c("No cannabis use screening", "Yes cannabis use screening")),
         hpdrgtalk = factor(ifelse(hpdrgtalk %in% c(85, 94, 97, 98, 99), NA, 
                                  ifelse(hpdrgtalk %in% c(2, 91, 93), 0, hpdrgtalk)), 
                           labels = c("No cannabis use talk", "Yes cannabis use talk")), 
         sutoutmrjpy = factor(ifelse(sutoutmrjpy == -9, NA, sutoutmrjpy), 
                              labels = c("No cannabis tx", "Yes cannabis tx")), 
         sutoutunsppy = factor(ifelse(sutoutunsppy == -9, NA, sutoutunsppy), 
                              labels = c("No unspecified drug tx", "Yes unspecified drug tx")), 
         cud_trtment = factor(ifelse(sutoutmrjpy == "Yes cannabis tx" | 
                                       sutoutunsppy == "Yes unspecified drug tx", 1, 0), 
                              labels = c("No treatment", "Yes treatment"))) |>
  as_survey_design(strata = vestr_c, 
                   ids = verep, 
                   weights = analwt2_c3, 
                   nest = T) 

# Subset to those with CUD 
cud_design <- nsduh_design |>
  filter(cud == "Yes CUD")


```


```{r}
nsduh_design |>
  group_by(cud) |>
  summarise(n = n(), 
            w_n = survey_total(),
            prop = survey_prop(vartype = "ci", proportion = T)) |>
  mutate(w_n = round(w_n)) |>
  mutate(across(c("prop", "prop_low", "prop_upp")) * 100)
  
```


In our sample (NSDUH 2021-2023), 6.45% (6.22 - 6.69) of the US population 12 and older had cannabis use disorder, which represent 18,175,923 individuals.

## Creating the Steps of the Cascade of Care


**STEP 1:** Defining the denominator--those with CUD (n = 14,427)
```{r}
step_1 <- cud_design |> 
  group_by(cud) |>
  summarise(n = n(), 
            prop = survey_prop(vartype = "ci", proportion = T), .groups = "drop") |>
  mutate(across(c("prop", "prop_low", "prop_upp")) * 100)

step_1 |>
  mutate(across(c("prop", "prop_low", "prop_upp")), round(2))

```


**STEP 2:** Among those 14,427 individuals (unweighted n) with CUD, what is the proportion that utilize healthcare in the past year?

Note: this variable has ~2.5% missings.
```{r}
step_2 <- cud_design |> 
  filter(!is.na(hlth_util)) |>
  group_by(cud, hlth_util) |>
  summarise(n = n(), 
            prop = survey_prop(vartype = "ci", proportion = T), .groups = "drop") |>
  mutate(across(c("prop", "prop_low", "prop_upp")) * 100)

step_2
```

**STEP 3:** Using the same 14,427 individuals with CUD, how many had cannabis screening? 

Note this variable had ~24% missings. For now, we are ignoring them, but we should probably address this.
```{r}
step_3 <- cud_design |> 
  filter(!is.na(hpusedrg)) |>
  group_by(cud, hpusedrg) |>
  summarise(n = n(), 
            prop = survey_prop(vartype = "ci", proportion = T), .groups = "drop") |>
  mutate(across(c("prop", "prop_low", "prop_upp")) * 100)

step_3
```

**STEP 4:** using the same denominator, how many receive treatment?

Note this variable had ~30% missings. For now, we are ignoring them, but we should probably address this.
```{r}
step_4 <- cud_design |> 
  filter(!is.na(cud_trtment)) |>
  group_by(cud, cud_trtment) |>
  summarise(n = n(), 
            prop = survey_prop(vartype = "ci", proportion = T), .groups = "drop") |>
  mutate(across(c("prop", "prop_low", "prop_upp")) * 100)

step_4
```

## Figure 1. Cascade of Care of Cannabis Use Disorder, NSDUH 2021-2023 (n=14,423)
```{r}
### first, I need to join all the steps into one single table to create the figure.
cascade_fig <- bind_rows(
  step_1 |>
  rename(step = cud) |>
  mutate(prop_low = NA, 
         prop_upp = NA),
  
  step_2 |>
  filter(hlth_util == "Yes Healthcare Utilization") |>
  rename(step = hlth_util) |>
  select(-cud),
  
  step_3 |>
  filter(hpusedrg == "Yes cannabis use screening") |>
  rename(step = hpusedrg) |>
  select(-cud),
  
  step_4 |>
  filter(cud_trtment  == "Yes treatment") |>
  rename(step = cud_trtment) |>
  select(-cud)
)


## changing the labels of the variable step
cascade_fig$step <- droplevels(cascade_fig$step)


cascade_fig <- cascade_fig |>
  mutate(step = fct_recode(step, 
                           "CUD" = "Yes CUD", 
                           "Used Healthcare" = "Yes Healthcare Utilization",  
                           "Screened" = "Yes cannabis use screening", 
                           "Treated" = "Yes treatment"))

## create the figure
cascade_plot <- cascade_fig |>
  ggplot(aes(x = step, y = prop, 
             ymin = prop_low, 
             ymax = prop_upp, 
             colour = step, 
             fill = step)) +
  geom_col() +
  geom_errorbar(color = "black",
                width = 0.2,
                linewidth = 0.5) +
  ggthemes::theme_clean() + 
  theme(legend.position = "none") + 
  labs(y = "%", 
       x = "", 
       caption = c("This cascade uses the first bar as the common denominator for each of the following steps")) + 
  scale_y_continuous(breaks = seq(0, 100, by = 10), 
                     expand = expansion(mult = c(0, 0.03)))
  
  
cascade_plot

```

